Embeddings and Attention in Predictive Modeling

### 2021-04-08 manuscript/manuscript.Rmd

In Table \@ref(tab:cv-results-attn), we show the cross-validated performance metrics of each of the models, as well as the two best performing models from the previous section. As before, we report both the root mean squared error (RMSE) and the mean absolute error (MAE). The results show that Model 5, which simply adds attention to Model 4, outperforms Model 4 on both the RMSE and MAE metrics. Thus, there is some room to argue that for tabular data modeling that uses embeddings for categorical data, the use of attention within models should be considered. Disappointingly, the TabNet architecture performs poorly, perhaps due to insufficient hyperparameter tuning, but indicating that \`\`out of the box", TabNet does not appear to work well for this modeling task. Finally, the TabTransformer model shows the best performance of all models considered in this research, beating the other models by a significant margin on both the RMSE and MAE metrics.

@@

Here, we explore what the TabTransformer model has learned from the data via the attention process that forms the core of the model. We focus on the difference between the flood embedding learned by the model, compared with the output of the Transformer layer that processes the embeddings before these enter the model. @Huang2020 call the latter \`\`contextual embeddings", because the attention mechanism within the TabTransformer model augments the embeddings based on the other covariates that appear for each observation $i$.

>>>>

In Table \@ref(tab:cv-results-attn), we show the cross-validated performance metrics of each of the models, as well as the two best performing models from the previous section. As before, we report both the root mean squared error (RMSE) and the mean absolute error (MAE). The results show that Model 5, which simply adds attention to Model 4, outperforms Model 4 on both the RMSE and MAE metrics. Thus, there is some room to argue that for tabular data modeling that uses embeddings for categorical data, the use of attention within models should be considered. The TabNet architecture performs poorly, perhaps due to insufficient hyperparameter tuning, but indicating that *out of the box*, TabNet does not appear to work well for this modeling task. Finally, the TabTransformer model shows the best performance of all models considered in this research, beating the other models by a significant margin on both the RMSE and MAE metrics.

@@

Here, we explore what the TabTransformer model has learned from the data via the attention process that forms the core of the model. We focus on the difference between the flood embedding learned by the model, compared with the output of the Transformer layer that processes the embeddings before these enter the model. @Huang2020 call the latter *contextual embeddings*, because the attention mechanism within the TabTransformer model augments the embeddings based on the other covariates that appear for each observation $i$.
